Compositional equivalences based on Open pNets

### 2021-10-21 2016-Weak-FH-Bisim/WeakBisim.tex
Minor

The parametric nature of the automata entails a second original aspect of FH-bisimulation: the nature of the bisimulation relation itself.
 A classical relation between states can be seen as a function mapping pairs of state to a boolean value (true if the states are related, false if they are not). An FH-bisimulation relation maps pairs of states to boolean expressions that use variables of the two systems.  %Each variable used in $\Pred_{s,t}$ should refer to a state variable of $s$ or $t$ and there must be no ambiguity on the referenced variable, we thus require that the two compared automata have disjoint state variables.
Formally, a relation over the states of two open automata  $\mylangle J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1 \myrangle$ and $\mylangle J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2 \myrangle$ has the signature $\mathcal{S}_1\times\mathcal{S}_2\to\AlgB$. 
We suppose without loss of generality that the variables of the two open automata are disjoint.
We adopt a notation similar to standard relations and denote it
 $\mathcal{R}=\{(s,t|\Pred_{s,t})\}$, where: 1) For any pair $(s,t)\in \mathcal{S}_1\times \mathcal{S}_2$, there is a 
   single
      $(s,t|\Pred_{s,t})\in\mathcal{R}$  stating that $s$ and $t$ are related 
      if $\Pred_{s,t}$       is 
      True, i.e. the states are related when the value of the automata variables% in $s$ and $t$
  verify the predicate $\Pred_{s,t}$. 2) The free variables of  $\Pred_{s,t}$ belong to $V_1$ and $V_2$, i.e. $\vars(\Pred_{s,t})\subseteq V_1\cup V_2$.
 FH-bisimulation is defined formally\footnote{In this article, we denote $\beta_{j x}$  a double indexed set, instead of the classical $\beta_{j,\, x}$. Indeed the standard notation would be too heavy in our case.}: 
 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim} ~\\
\noindent
\begin{minipage}{0.67\linewidth} 	Suppose 
   $A_1 = \mylangle J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1 \myrangle$ and $A_2 = \mylangle J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2 \myrangle$
   are open automata with identical holes of the same sort, with disjoint sets of variables ($V_1\cap V_2=\emptyset$).

 Then 
$\mathcal{R}$ is an FH-bisimulation if and only if for any  states
$s\in\mathcal{S}_1$ and $t\in\mathcal{S}_2$, $(s,t|\Pred_{s,t})\in\mathcal{R}$, we 
have
the following:
\end{minipage}
\hspace{2mm}
\begin{minipage}{0.30\linewidth}
	\includegraphics[width=\linewidth]{XFIG/Bisim}
\end{minipage}

\end{mathpar}
 there exists an indexed set of  open transitions $OT_x^{x\in X} \subseteq \mathcal{T}_2$:
 \begin{mathpar}
%    \left( fresh \ \set{\alpha_i}, \set{b_j}, v_x.\ \
    \openrule
         {
           \beta_{j x}^{j\in J_{x}}, \Pred_{OT_x},\Post_{OT_x}}
         {t \OTarrow {\alpha_x} t_x}
%         \right)
\end{mathpar}
 such that  $\forall x.\, J'=J_{x}$ and there exists $\Pred_{s',t_x}$ such that $(s',t_x|\Pred_{s',t_x})\in 
 \mathcal{R}$
 and  
\begin{multline*}
 \Pred_{s,t} \land \Pred_{OT}\implies\\
%\hspace{1cm}
 \displaystyle{\bigvee_{x\in X}
   \left( \forall j. \beta_j=\beta_{jx}  \land \Pred_{OT_x}
     \land \alpha\!=\!\alpha_x \land  
     \Pred_{s',t_x}\subst{\Post_{OT}\!\uplus\!\Post_{OT_x}}\right)}
\end{multline*}
     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
     %     \right)$.
%\bigskip
%
% $\Pred_{s,t} \land \Pred_{OT}\implies\bigvee_{x\in X} \Pred_x$
%and\\
%%\hspace{1cm}
% $ \forall{x\in X}. \Pred_x \land \Pred_{s,t} \land \Pred_{OT} \Rightarrow
%   \left( \forall j. \beta_j=\beta_{jx}  \land \Pred_{OT_x}
%     \land \alpha\!=\!\alpha_x \land  
%     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)$
%     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
%     %     \right)$.

@@

\subsubsection*{Predicates} 
%Let
%$\mylangle\set{\pNet},\set{\Sort},\set{\symb{SV}}\myrangle$
%be a pNet. 
 We 
define a
predicate $\Predsv$ relating a synchronisation vector (of the form $\SV{{(\alpha'_i)}^{i\in I}, {(\beta'_j)}^{j\in J}} 
{\alpha'} 
{e_b}$),
the actions of the involved sub-pNets and the resulting actions.

 This predicate verifies:
\begin{multline*}
\Predsv \Big(\left(\SV{{(\alpha'_i)}^{i\in I}, {(\beta'_j)}^{j\in J}} 
{\alpha'} 
{e_b}\right), \alpha_i^{i\in I}, \beta_j^{j\in J}, \alpha\Big)\Leftrightarrow \\~\qquad\qquad%\bigg(
%
%\exists {(\alpha'_i)}^{i\in I},
%{(\beta'_j)}^{j\in J},v'.\, SV=
%\\~~\land
\forall i\in I.\, \alpha_i=\alpha'_i\land \forall j \in J.\, \beta_j=\beta'_j \land 
\alpha=\alpha' 
\land e_b
\end{multline*}

>>>>

The parametric nature of the automata entails a second original aspect of FH-bisimulation: the nature of the bisimulation relation itself.
 A classical relation between states can be seen as a function mapping pairs of state to a boolean value (true if the states are related, false if they are not). An FH-bisimulation relation maps pairs of states to boolean expressions that use variables of the two systems.  %Each variable used in $\Pred_{s,t}$ should refer to a state variable of $s$ or $t$ and there must be no ambiguity on the referenced variable, we thus require that the two compared automata have disjoint state variables.
Formally, a relation over the states of two open automata  $\mylangle J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1 \myrangle$ and $\mylangle J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2 \myrangle$ has the signature $\mathcal{S}_1\times\mathcal{S}_2\to\AlgB$. 
We suppose without loss of generality that the variables of the two open automata are disjoint.
We adopt a notation similar to standard relations and denote it
 $\mathcal{R}=\{(s,t|\Pred_{s,t})\}$, where: 1) For any pair $(s,t)\in \mathcal{S}_1\times \mathcal{S}_2$, there is a 
   single
      $(s,t|\Pred_{s,t})\in\mathcal{R}$  stating that $s$ and $t$ are related 
      if $\Pred_{s,t}$       is 
      True, i.e. the states are related when the value of the automata variables% in $s$ and $t$
  verify the predicate $\Pred_{s,t}$. 2) The free variables of  $\Pred_{s,t}$ belong to $V_1$ and $V_2$, i.e. $\vars(\Pred_{s,t})\subseteq V_1\cup V_2$.
 FH-bisimulation is defined formally\footnote{In this article, we denote $\beta_{j x}$  a double indexed set, instead of the classical $\beta_{j,\, x}$. Indeed the standard notation would be too heavy in our case.}: 
 \begin{definition}[Strong FH-bisimulation]\label{def-FH-bisim} ~\\
\noindent
	Suppose 
   $A_1 = \mylangle J,\mathcal{S}_1, s_0,V_1,
   \mathcal{T}_1 \myrangle$ and $A_2 = \mylangle J,\mathcal{S}_2,t_0,V_2, \mathcal{T}_2 \myrangle$
   are open automata with identical holes of the same sort, with disjoint sets of variables ($V_1\cap V_2=\emptyset$).

 Then 
$\mathcal{R}$ is an FH-bisimulation if and only if for any  states
$s\in\mathcal{S}_1$ and $t\in\mathcal{S}_2$, $(s,t|\Pred_{s,t})\in\mathcal{R}$, we 
have
the following:

 \begin{itemize}
 \item   
For any open transition $OT$ in $\mathcal{T}_1$:\\
\begin{minipage}{0.67\linewidth}  \begin{mathpar}
     \openrule
         {
           \beta_j^{j\in J'},\Pred_{OT},\Post_{OT}}
         {s \OTarrow {\alpha} s'}

\end{mathpar}
 there exists an indexed set of  open transitions $OT_x^{x\in X} \subseteq \mathcal{T}_2$:
 \begin{mathpar}
%    \left( fresh \ \set{\alpha_i}, \set{b_j}, v_x.\ \
    \openrule
         {
           \beta_{j x}^{j\in J_{x}}, \Pred_{OT_x},\Post_{OT_x}}
         {t \OTarrow {\alpha_x} t_x}
%         \right)
\end{mathpar}

\end{minipage}
\hspace{2mm}
\begin{minipage}{0.30\linewidth}
\vspace{-2em}
{	\includegraphics[width=\linewidth]{XFIG/Bisim}}
\end{minipage}
 such that  $\forall x.\, J'=J_{x}$ and there exists $\Pred_{s',t_x}$ such that $(s',t_x|\Pred_{s',t_x})\in 
 \mathcal{R}$
 and  
\begin{multline*}
 \Pred_{s,t} \land \Pred_{OT}\implies\\
%\hspace{1cm}
 \displaystyle{\bigvee_{x\in X}
   \left( \forall j. \beta_j=\beta_{jx}  \land \Pred_{OT_x}
     \land \alpha\!=\!\alpha_x \land  
     \Pred_{s',t_x}\subst{\Post_{OT}\!\uplus\!\Post_{OT_x}}\right)}
\end{multline*}
     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
     %     \right)$.
%\bigskip
%
% $\Pred_{s,t} \land \Pred_{OT}\implies\bigvee_{x\in X} \Pred_x$
%and\\
%%\hspace{1cm}
% $ \forall{x\in X}. \Pred_x \land \Pred_{s,t} \land \Pred_{OT} \Rightarrow
%   \left( \forall j. \beta_j=\beta_{jx}  \land \Pred_{OT_x}
%     \land \alpha\!=\!\alpha_x \land  
%     \Pred_{s',t_x}\subst{\Post_{OT}\uplus\Post_{OT_x}}\right)$
%     %     \symb{Subst}(\Pred_{target_x}, \Post_{OT} o \Post_{OT_x})
%     %     \right)$.

@@

\subsubsection*{Predicates} 
%Let
%$\mylangle\set{\pNet},\set{\Sort},\set{\symb{SV}}\myrangle$
%be a pNet. 
 We 
define a
predicate $\Predsv$ relating a synchronisation vector (of the form $\SV{{(\alpha'_i)}^{i\in I}, {(\beta'_j)}^{j\in J}}{\alpha'} {e_b}$),
the actions of the involved sub-pNets and the resulting actions.

 This predicate verifies:
\begin{multline*}
\Predsv \Big(
\big({\SV{{(\alpha'_i)}^{i\in I}, {(\beta'_j)}^{j\in J}}{\alpha'}{e_b}}\big)
, \alpha_i^{i\in I}, \beta_j^{j\in J}, \alpha\Big)\Leftrightarrow \\
\forall i\in I.\, \alpha_i=\alpha'_i\land \forall j \in J.\, \beta_j=\beta'_j \land 
\alpha=\alpha' 
\land e_b
\end{multline*}
